PANCHANGE_data_preprocessing

### 2020-08-13 PANCHANGE_data_preprocessing2020-07-15.Rmd
read in RMAP employment data

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = TRUE
RAMP = FALSE
```

@@

  nbr.dem.raw.id <- nbr.dem.raw %>%
    drop_na(subjectid) %>% # Drop NAs
    distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
    separate(ID.intm, into = "ID", sep = 7) %>%
    mutate(ID = as.numeric(ID),
           startDate.coping = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.coping, #Date when participant signed up to nbr
#      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Birthyear = demographics.year, # Birthyear
      Gender = demographics.which_gender_do_you_identify_with, # Gender
      Sex = demographics.select_questionnaire_items_medical, #Sex
      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

# Create categorical age groups per 10 years
dem.raw.id <- dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )

dem.raw.id <- dem.raw.id %>%
    mutate(
      age_category =
        recode_factor(
          age_category_numeric, 
          "1" = "16 to 18 years",
          "2" = "19 to 25 years",
          "3" = "26 to 35 years",
          "4" = "36 to 45 years",
          "5" = "46 to 55 years",
          "6" = "56 to 65 years",
          "7" = "66 to 70 years",
          "8" = "71 to 75 years",
          "9" = "76 to 80 years",
          "10" = "81 to 85 years",
          "11" = "86 to 90 years",
          "12" = "91 to 100 years",
          "13" = "Above 100 years"
          )
    )

dem.raw.id %>%
  freq(age_category)

}
```

>>>>

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = FALSE
RAMP = TRUE
```

@@

  nbr.dem.raw.id <- nbr.dem.raw %>%
    drop_na(subjectid) %>% # Drop NAs
    distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
    separate(ID.intm, into = "ID", sep = 7) %>%
    mutate(ID = as.numeric(ID),
           startDate.coping = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.coping, #Date when participant signed up to nbr
#      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Birthyear = demographics.year, # Birthyear
      Gender = demographics.which_gender_do_you_identify_with, # Gender
      Sex = demographics.select_questionnaire_items_medical, #Sex
      EduYrs = demographics.school_education_school_years.txt, # Years in education
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

# Create categorical age groups per 10 years
dem.raw.id <- dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )
}

dem.raw.id <- dem.raw.id %>%
    mutate(
      age_category_full =
        recode_factor(
          age_category_numeric, 
          "1" = "16 to 18 years",
          "2" = "19 to 25 years",
          "3" = "26 to 35 years",
          "4" = "36 to 45 years",
          "5" = "46 to 55 years",
          "6" = "56 to 65 years",
          "7" = "66 to 70 years",
          "8" = "71 to 75 years",
          "9" = "76 to 80 years",
          "10" = "81 to 85 years",
          "11" = "86 to 90 years",
          "12" = "91 to 100 years",
          "13" = "Above 100 years"
          )
    )

dem.raw.id %>%
  freq(age_category_full)

# RAMP employment
## Read in data
```{r ramp emp cols}
if(RAMP == TRUE){
  ramp.emp.raw <- readRDS(file = paste0(data.raw_path, "/ramp/employment_ramp.rds"))
  dim(ramp.emp.raw)
  colnames(ramp.emp.raw)
}
```

# Select variables
```{r ramp emp cleaning}
if(RAMP == TRUE){
  exclude_cols_emp <- c("ID",
                        "Sample")

ramp.emp.raw$externalDataReference <- ramp.emp.raw$`Login ID` #move Login ID to external data reference

ramp.emp.raw.id <- ramp.emp.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.ramp = startDate) %>% # Rename startDate
    select(
      highest_education_ramp = employment.what_is_your_highest_level_of_education,
      # EduYrs = employment.university_education_attend_school.txt, # Years in education - not in RAMP dataset
      employment.government_work_key_workers
    ) %>%
    add_numeric(., exclude = exclude_cols_emp) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(ramp.emp.raw.id)
  # Inspect colnames
  colnames(ramp.emp.raw.id)
  #Differences
  dim(ramp.emp.raw)[1]-dim(ramp.emp.raw.id)[1]
}
```

```{r ramp copy emp}
if(RAMP == TRUE){
  emp.raw.id <- ramp.emp.raw.id
}
```
